---
title: PA1_template.Rmd
author: "Kyaw Thu"
date: "Monday, April 13, 2015"
output: html_document
---

#**Loading and preprocessing the data**#

```{r setoptions, eval=FALSE}
opts_chunk$set(echo=TRUE,results="asis")
cache=TRUE

```{r}
activity <- read.csv("activity.csv", header=TRUE)
activity$date = as.Date(activity$date, format="%Y-%m-%d")
names(activity)

```


# Remove all NA#
```{r}
na_rm_activity<-activity[!is.na(activity$steps),]

```

**What is mean total number of steps taken per day**#

```{r}
Totalsteps <- tapply(na_rm_activity$steps, na_rm_activity$date, sum)
```

**Total number of steps taken per day and ignore NA value**
```{r}
Totalsteps<-Totalsteps[!is.na(Totalsteps)]


```

#Plotting histogram of total number of steps taken each day
par("mar"=c(5.1, 4, 4.1, 2.1))

```{r hist,fig.height=4}
hist(Totalsteps,col='orange',main="Histogram of Total steps taken per day")
```

# Mean and median of steps taken each day

```{r}

mean(Totalsteps)
median(Totalsteps)
```
#**What is the average daily activity pattern?**#

Average daily activity pattern can be analysed by timeseries function and produced its plot.

Daily activity pattern timeseries
```{r}
DailyPattern <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
```

Histogram of Timeseries
```{r}
plot(row.names(DailyPattern), DailyPattern, type = "l", xlab = "5-min interval", ylab = "Average across all Days", main = "Time Series Analysis\n Average   daily activity pattern", col = "blue")
```

**Maximum daily steps**
5-minute interval, on average across all the days in the dataset, contains the maximum number of steps 835.


5-minute interval, on average across all the days in the dataset, contains the maximum number of steps 835.

```{r}
which.max(DailyPattern)
```

#**Imputing missing values**#

**Report of the total number of missing values in the dataset**

*2304 NA values included in activity dataset.*

```{r}
sum(is.na(activity))
```

**Filling in all of the missing values in the dataset**
```{r}
intervals <- unique(activity$interval)
new_activity <- activity
for (int in intervals) {
  num_missing <- nrow(new_activity[new_activity$interval == int & is.na(new_activity$steps),])
  if (num_missing > 0) {
    new_activity[new_activity$interval == int & is.na(new_activity$steps),]$steps <- rep(DailyPattern[as.character(int)], num_missing)
  }
}

new_steps <- tapply(new_activity$steps, new_activity$date, sum)
```
**New Plot**
```{r}

hist(new_steps, col='green',main='Histogram of total steps taken per day \n New figure')

```
**Report on mean and median of steps taken each day**

It was found that both mean and mean are the same value of 10766 steps. 

```{r}
mean(new_steps)

median(new_steps)

```

Results of mean and median values are essentially the same.

**Impact on imputing NA values in making estimate**
It was discovered that there is only negligable difference in the mean and median values from imputing missing NA values from the dataset.

#**Are there differences in activity patterns between weekdays and weekebds?**#

**Create a new factor variable in the dataset with two levels - "weekday" and "weekend"**

```{r}
days <- weekdays(activity$date)
daylevel <- vector()
  for (i in 1:nrow(activity)) {
    if (days[i] == "Saturday") {
        daylevel[i] <- "Weekend"
    } else if (days[i] == "Sunday") {
        daylevel[i] <- "Weekend"
    } else {
        daylevel[i] <- "Weekday"
    }
}

activity$daylevel <- daylevel
activity$daylevel <- factor(activity$daylevel)
```

**Weekday vs. weekend activity**
The difference between weekend and weekdays values can be visualized in the following graph:

```{r}
library(lattice)

AvgSteps <- aggregate(steps ~ interval + daylevel, data = activity, mean)
names(AvgSteps) <- c("interval", "daylevel", "steps")
xyplot(steps ~ interval | daylevel, AvgSteps, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")
```


